---
title: "automated_reporting"
author: "Tian Yu"
date: "9/23/2020"
output: html_document
---

```{r}
library(tidyverse)
library(officer)
library(flextable)
```

# PowerPint Slide Master

## Importing the Template 

```{r 1.1 importing the template}
my_file <- file.path("input","templates","Tutorial Template.pptx") %>% 
                        read_pptx()

class(my_file) # checking if correct class

my_file %>%
  layout_summary()
```

## Creating a PowerPoint Deck

```{r 1.2 creating a PPT deck}
pptx_obj <- read_pptx () # new empty file
pptx_obj <- pptx_obj %>% 
  add_slide(layout  = 'Title and Content',master = "Office Theme")

pptx_obj %>% print(target = "output/first_example.pptx")
```

We can not load the themes of Office *ex nihilo* returns error
```{r themes1}
pptx_obj <- read_pptx() # new empty file
pptx_obj <- pptx_obj %>% 
  add_slide(layout = "Title and Content", master = "Integral")
```

However, we can save an empty pptx with the desired theme and use it as a template
```{r themes1}
pptx_obj <- read_pptx(file.path("input","templates","integral.pptx"))
layout_summary(pptx_obj)
```

We can even load a template with more than one theme
```{r themes2}
pptx_obj <- read_pptx(file.path("input","templates","multmasters.pptx"))
layout_summary(pptx_obj)
```

## Selection Pane

## Key functions: read_pptx(path), layout_summary(x), layout_properties(x), add_slide(x, layout, master), on_slide(x, index), slide_summary(x, index = NULL)


## Example Code
```{r 1.5 example code}
pptx_obj <- read_pptx() # new empty file

pptx_obj <- pptx_obj %>% # add slide
  add_slide(layout = "Title and Content", master = "Office Theme")

layout_summary(pptx_obj) # contains only basic layouts
layout_properties(pptx_obj) # additional detail

pptx_obj <- pptx_obj %>% 
  on_slide(index = 1) # set active slide
slide_summary(pptx_obj) # slide is empty
```

# Placeholders

## Placeholders and Shapes

```{r 2.1 shapes1}
# Example 1
my_data <- c("My functions are:","ph_with","ph_location_type")
my_type <- "body"

pptx_obj <- read_pptx() # new empty file

pptx_obj <- pptx_obj %>% 
  add_slide(layout = "Title and Content", master = "Office Theme")

pptx_obj <- pptx_obj %>%
  ph_with(value = my_data,location = ph_location_type(type = my_type))

pptx_obj %>%
  print(target = "output/test2.1.pptx") 
```

```{r 2.2 shapes2}
# Example 2: 
my_data <- head(mtcars)[,1:4]
my_type <- "body"

pptx_obj <- read_pptx() # new empty file
pptx_obj <- pptx_obj %>% 
    add_slide(layout = "Title and Content", master = "Office Theme")

pptx_obj <- pptx_obj %>% 
  ph_with(value = my_data, location = ph_location_type(type = my_type)) 

pptx_obj %>%
  print(target = "output/test2.2.pptx") 
```

```{r 2.3 shapes3}
# Example 3
# We add a text box item in a custom position
# The same can be done for an image, logo, custom objects, etc.

my_data <- "My text"
my_type <- "body"

pptx_obj <- read_pptx() # new empty file
pptx_obj <- pptx_obj %>%
  add_slide(layout = "Title and Content", master = "Office Theme")

#ph_location is a subfunction which takes as argument 
#left/top/width/height, units are inches

pptx_obj <- pptx_obj %>%
  ph_with(value = my_data, location = ph_location(left = 2, top = 2, width = 3, height = 1))

pptx_obj %>%
  print(target = "output/test2.3.pptx") 
```

## Key functions: ph_with()


# Text

## Working with Text
Each new text item added to a PowerPoint via officer is a paragraph object
fpar() (“formatted paragraph”) creates this object

```{r 3.1 text1}
my_data <- fpar("My text")
my_type <- "body"

pptx_obj <- read_pptx() # new empty file
pptx_obj <- pptx_obj %>% 
  add_slide(layout = "Title and Content", master = "Office Theme")

## Add paragraph
pptx_obj <- pptx_obj %>% 
  ph_with(value = my_data, location = ph_location_type(type = my_type))

## Try to add a second paragraph
my_data2 <- fpar("My other text")
my_type <- "body"

pptx_obj <- pptx_obj %>% 
  ph_with(value = my_data2, location = ph_location_type(type = my_type) )

pptx_obj %>%
  print(target = "output/test3.1.pptx") 
## As we see, this code doesn’t produce bullet points as we might hope
```

block_list() allows us to wrap multiple paragraphs together

```{r 3.2 text2 blocklist}
my_data <- fpar("My text")
blank_line <- fpar("")
my_data2 <- fpar("My other text")

my_list <- block_list(my_data, blank_line, my_data2)

my_type <- "body"

pptx_obj <- read_pptx() # new empty file
pptx_obj <- pptx_obj %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = my_list, location = ph_location_type(type = my_type) )

pptx_obj %>%
  print(target = "output/test3.2.pptx") 
```

Use ftext() (“formatted text”) to edit the text before pasting into paragraphs. ftext() requires a second argument called prop which contains the formatting properties.
```{r 3.3 text3}
my_prop <- fp_text(color = "red", font.size = 16)
my_text <- ftext("Hello", prop = my_prop)

my_par <- fpar(my_text) ## formatted
blank_line <- fpar("")

my_par2 <- fpar("My other text") ## unformatted
my_list <- block_list(my_par,blank_line, my_par2)

pptx_obj <- read_pptx() # new empty file

pptx_obj <- pptx_obj %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = my_list, location = ph_location_type(type = my_type) )

pptx_obj %>%
  print(target = "output/test3.3.pptx") 
```

## Key functions: fpar(), ftext(), fp_text(), block_list()

## Example Code
```{r 3.4 example code}
my_list <- block_list(
  fpar(ftext("Hello", prop = fp_text(color = "red", font.size = 16))) ,
  fpar(ftext("World", prop = fp_text(color = "blue", font.size = 14))) )

# The hierarchy is:
# block_list > fpar > ftext > fp_text

pptx_obj <- read_pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = block_list(
    fpar(ftext("Hello", prop = fp_text(color = "red", font.size = 16))) ,
    fpar(ftext("World", prop = fp_text(color = "blue", font.size = 14)))),
    ph_location_type(type = "body")
  ) %>%
  print(target = "output/test3.4.pptx") 
```

# Tables
## Basic Code
```{r 4.1 basic code}
my_data <- head(mtcars)

pptx_obj <- read_pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = head(mtcars), location = ph_location_type(type = "body")) %>%
  print(target = "output/test4.1.pptx") 
```

## Introduction to flextable
```{r 4.2 flextable1}
base_table <- tutorial_data[["Age"]] %>%
  table() %>%
  enframe(name = "Age", value = "Count") %>%
  mutate(Count = as.integer(Count)) %>%
  mutate(Percent = format_percent(Count / sum(Count), num_decimals = 2))

base_table 
```

flextable: create attractive tables with predefined formatting. Use of %>% is recommended for readability
```{r flextable2}
library(flextable)

ft_table <- base_table %>%
  flextable()

ft_table # see preliminary result in Viewer tab of RStudio
```

## Demonstration

```{r 4.4 demonstration}
ft_table <- base_table %>%
  flextable() %>%
  autofit() %>% # straightforward, column width
  # ALIGNMENT
  # alignment of header: we use part argument
  align(align = "center", part = "header") %>%
  # alignment of content: we can use part = "body" or specify exact lines 
  align(i = 1:nrow(base_table), j = 1:ncol(base_table), align = "center") 

ft_table
```

Set font and characters
```{r 4.5 demonstration cont.1}
ft_table = ft_table %>% 
# FONT AND CHARACTERS
# each command is independent, there are no nested functions as in officer
bold(i = 1, j = 1:3) %>% # first row, all cols
  italic(i = 3, j = ~Age+Count+Percent) %>% # first row, all cols, using ~ notation
  
  fontsize(i = 2, size = 16) %>%
  font(fontname = "Calibri") %>% # since no i or j are input, change is for all data
  font(fontname = "Roboto", part = "header") %>% #different font for header
  color(i = 3, j = 2, color = "red") %>%
  
  # WIDTH AND HEIGHT
  # all measurements are in inches
  width(j = 1, width = 4) %>% # column 1 wider
  height(i = 8, height = 0.5) %>% # row 8 change
  # CELL COLORS (background) 
  bg(bg = "#0088FF", part = "header") %>% # a custom background for the header
  bg(i = 7:10, bg = "#C5C5C5") # a custom background for some cells

ft_table
```

Set borders
```{r 4.6 demonstration cont.2}
#BORDERS
# For borders we need to use nested functions (similar to fpar>ftext>fp_text)
#fp_border() is the second level function we will use to specify border"s characteristics
# as argument it takes color, style, and width
  
my_border <- fp_border(color = "black", style = "solid", width = 2)

# We use this second level function inside various main border functions
# border_outer()
# border_inner()
# border_inner_h()
# border_inner_v()
ft_table <- ft_table %>%
  border_outer(part = "all", border = my_border) %>% # using predefined border
  border_inner(part = "body", border = fp_border(style = "dashed"))

ft_table
```

## Demonstration Output

## Reformat original example

```{r 4.7 original example}
## ft parameters
header_background_color = a_green
body_background_color = a_cream
header_text_col = a_cream
body_text_col = a_dark_grey
total_text_col = a_red

border_solid <- fp_border(color = a_dark_grey, width = 2)
border_dashed <- fp_border(color = a_dark_grey, width = 1, style = "dashed")

ft <- summary_table %>%
  flextable() %>%
  font(fontname = font_name, part = "all") %>%
  fontsize(size = font_size, part = "all") %>%
  bold(part = "header") %>%
  border_remove() %>%
  border_outer(border = border_solid) %>%
  border_inner_v(border = border_solid, part = "header") %>%
  border_inner_h(border = border_dashed, part = "header") %>%
  border_inner_v(border = border_solid, part = "body") %>%
  border_inner_h(border = border_dashed, part = "body") %>%
  bg(bg = header_background_color, part = "header") %>%
  bg(bg = body_background_color, part = "body") %>%
  align(align = "center", part = "all") %>%
  color(color = header_text_col, part = "header") %>%
  color(color = body_text_col, part = "body") %>%
  color(color = total_text_col, part = "body", i = nrow(summary_table)) %>%
  italic(part = "body", i = nrow(summary_table)) %>%
  bold(part = "body", i = nrow(summary_table)) %>%
  autofit(add_w = 1)

ft
# Add table to slide
pptx_obj <- read_pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = ft, ph_location(left = 2, top = 2, width = 4 )) %>% 
  print(target = "output/test4.7.pptx") 
# positions are fixed. We can find exact positions to center the table

```
## key functions: flextable(), align(), bold(), font(), color(), bg(), height() & width(), border_outer() & border_inner() & border_inner_h() & border_inner_v(), autofit()

Additional function to learn: merge(), compose() & as_chunk(), style()

fix_border_issues()

# Charts

## Adding charts as images

```{r 5.1 rvg package}
# We have preloaded a function to plot the chart.
# the function is using ggplot2 as plotting library

chart_to_plot <- sample_data_list[['Sample 1']] %>%
  make_jar_chart() # code to create a ggplot2 item, we will skip the contents
print(chart_to_plot) # see in Plots Window
```

## rvg Example

```{r 5.2 rvg example}
# the output is a ggplot2 object

# To add this object as a rvg object on a slide, we will use the ph_with_vg
# ph_with_vg replaces the ph_with for a rvg object
# ph_with_vg_at allows to input a precise position for a chart, using the top/left we know already
# all units are in inches
# argument code requires print(chart), argument type is a specific place on slide ("body" or other)


## IMPORTANT: ph_with_vg is is deprecated.
#old syntaxis ph_with_vg(code = print(chart_to_plot), type = "body") 

pptx_obj <- read_pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = chart_to_plot), location  =  ph_location_type(type = 'body')) %>%
  print(target = "output/5.2.rvg1.pptx") 

## IMPORTANT: ph_with_vg_at is is deprecated.
# old syntaxis ph_with_vg_at(code = print(chart_to_plot), left = 1, top = 1, width = 8, height = 6) 

pptx_obj <- read_pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = dml(ggobj = chart_to_plot), location =  ph_location(left = 1, top = 1, width = 8, height = 6))  %>%
  print(target = "output/5.2.rvg2.pptx") 

# all items on the chart inside the pptx are now editable, just click on any and see 
# the Shape Format tab in PowerPoint
```

## mschart Package

```{r 5.3 mschart example}
# sample dataframe
mydata <- sample_data_list[['Sample 1']] %>%
  group_by(Variable,Response) %>% count()

# syntaxis is similar to ggplot2"s aes() with x,y,group
my_barchart <- ms_barchart(data = mydata, x = "Variable", y = "n", group = "Response")

# to add the object to a powerpoint slide we can use the officer"s native ph_with
pptx_obj <- read_pptx() %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>%
  ph_with(value = my_barchart, location = ph_location_type(type = "body")) %>%
  print(target = "output/5.3.msoffice.pptx") 

# if we would open a rvg slide and an ms office slide and click on the slide
# for the rvg slide the only menu that appear is shape format
# While for the msoffice chart we have now the chart design option with all msoffice functionalities
# by using chart_settings() functions one can customise in R the charts
```

# Word

## Word documents
Formats share similarities

body_add_*()

```{r 6.1 word doc1}
my_doc <- read_docx() %>% 
  body_add_par(value = "My Text", style = "Normal") %>%
  body_add_par(value = "Other Text", style = "Normal") %>% 
  body_add_par(value = "Conclusion", style = "Normal") %>% 
  print(target = 'output/6.1.mydoc.docx')
```

body_add_break()

```{r 6.2 word doc2}
my_doc <- read_docx() %>% 
  body_add_par(value = "My Text", style = "Normal") %>%
  body_add_break() %>% 
  body_add_par(value = "Conclusion", style = "Normal") %>% 
  print(target = 'output/6.2.mydoc.docx')
```

```{r 6.3 word doc3}
my_format <- fp_text(font.family = 'Calibri', font.size = 14, bold = TRUE, color = 'red')
my_text <- ftext('My dataset is:', my_format)
my_par <- fpar(my_text)

doc <- read_docx() %>% 
  body_add_par(value = "Table of content", style = "heading 1") %>% 
  body_add_par(value = "", style = "Normal") %>% 
  body_add_fpar(my_par, style = "Normal") %>% #formatted paragraph function
  body_add_par(value = "", style = "Normal") %>% 
  body_add_table(value = head(mtcars)[, 1:4], style = "table_template" ) %>% 
  print(target = 'output/6.3.mydoc.docx')

read_docx() %>% styles_info() 
```

